@@ -1,4 +1,9 @@
 -# Reproducible Research: Peer Assessment 1 - CW Moellenkamp
 +---
 +title: "Reproducible Research: Peer Assessment 1"
 +output: 
 +  html_document:
 +    keep_md: true
 +---
  
  
## Loading and preprocessing the data 

*Read Table and create second version removing NAs for initial processing*

```{r}
stepsraw<-read.csv("activity.csv")
stepscomp<-stepsraw[complete.cases(stepsraw),]
```

##Report on stepscomp table data  

*Calculate the total number of steps per day*

```{r}
stepspd<-tapply(stepscomp$steps,stepscomp$date,sum)
stepsperday<-stepspd[complete.cases(stepspd)]
stepsperday
```

*Make a histogram of the number of steps per day*

```{r}
hist(stepsperday)
```

*Report the mean and median of the steps per day*

```{r}
meansteps<-mean(stepsperday)
mediansteps<-median(stepsperday)
meansteps
mediansteps
```

*Create time series plot of the 5 minute interval average steps*

```{r}
avintsteps<-tapply(stepscomp$steps,stepscomp$interval,mean)
Intervals<-names(avintsteps)
plot(Intervals,avintsteps,type="l",ylab="Average Steps per Interval")
```

*Report which interval on average has the maximum steps*

```{r}
maxstepinterval<-names(avintsteps[max(avintsteps)])
maxstepinterval
```

##Impute Missing Values in the original data file called stepsraw

*Report the total number of missing values in the dataset*

```{r}
NAvals<-sum(is.na(stepsraw$steps))
NAvals
```

*Develop strategy to replace NAs*  
**Use overall daily mean divided by 288 intervals per day to fill in missing data**

```{r}  
meaninterval<-meansteps/288
stepsimpute<-stepsraw
stepsimpute$steps[which(is.na(stepsimpute$steps))]<-meaninterval
```

*Create histogram of this new dataset stepsimpute on a steps per day basis*

```{r}
stepsimppd<-tapply(stepsimpute$steps,stepsimpute$date,sum)
hist(stepsimppd)
```

*Report mean and median of the imputed data set*

```{r}
meanstepsimp<-mean(stepsimppd)
medianstepsimp<-median(stepsimppd)
meanstepsimp
medianstepsimp
```

**These data differ by adding steps at a mean value across all intervals - the** **mean is slightly higher and median is the same as in the data with NAs removed since more steps were added when the mean per interval values were added**


**The impact of imputing missing data is pushing more steps into the center**
**of the histogram in the grouping containing the mean of 10,766 steps per day**

##Now lets add a factor noting if a date is a weekend or weekday  

*This will allow us to see when more steps are taken on average*

```{r}
library(dplyr)
wday<-c("Monday","Tuesday","Wednesday","Thursday","Friday")
stepsimpute<-mutate(stepsimpute,dayfactor=
                      factor(weekdays(as.Date(stepsimpute$date)) %in% wday,
                      levels=c(TRUE,FALSE),labels=c("DAY","END")))
```

*Create a plot of averages of weekdays in one panel and*
*weekends in another panel*

```{r}
stepsimputeend<-stepsimpute[stepsimpute$dayfactor=="END",]
stepsimputewday<-stepsimpute[stepsimpute$dayfactor=="DAY",]

endimpavstep<-tapply(stepsimputeend$steps,stepsimputeend$interval,mean)
Intervalsend<-names(endimpavstep)

dayimpavstep<-tapply(stepsimputewday$steps,stepsimputewday$interval,mean)
Intervalsday<-names(dayimpavstep)

par(mfrow=c(2,1))

plot(Intervalsday,dayimpavstep,type="l",ylab="Average Steps per Interval",
     xlab="Intervals", main ="Weekday")

plot(Intervalsend,endimpavstep,type="l",ylab="Average Steps per Interval",
     xlab="Intervals", main="Weekend")
```
